Thermodynamic fragments in magic state theories

### 2020-12-03 magic_paper.tex

\subsubsection*{Strange state distillation in the unital fragment}\label{sec:lcsu}

In this section we study the MSD process,
\begin{equation}\label{eq:sudist}
	\noisysn \xrightarrow{\E \in \O_{\id/3}} \noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{\mm},\ \nn = \tt + \mm,
\end{equation}
where we have uniformised the number of qutrits on both sides of the process.

In~\cref{app:lcsu_technical}, we define explicit expressions for the Lorenz curve of the $\NN$-copy, $\epsilon$-noisy Strange state in the unital fragment, 
\begin{equation}\label{eq:lcsu}
	\lc{\noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{\mm}}{\frac{1}{d}\id}(x),\ x\in[0,1].
\end{equation}
We call Lorenz curves of this type \emph{standard Lorenz curves}. \nick{expand (non-uniformity)}

Lorenz curves are fully defined by the coordinates of their elbows which is clear in~\cref{fig:lcs} for the Strange state in particular.
The exact expressions for the elbow coordinates of a standard Lorenz curve can take $2^3 = 8$ different forms, depending on whether the noise level $\epsilon$ is low/high, the parity of the number of copies $\NN$ is even/odd and the location of the elbow relative to the maximum is on the left hand side (LHS -- includes maximum) / right hand side (RHS).
They are given in full detail in~\cref{app:lcsu_coord_elb}.

Here we focus on the part of the process described by the LHS of the Lorenz curves, in low noise regimes ($0 \leq \epsilon < 3/7$) with even numbers of initial and target copies.
In this case, the elbow expressions of the standard Lorenz curve in~\cref{eq:lcsu} are given by
\begin{align}
	x_i &= \Phi_+\left(2\ii;\NN,\frac{8}{9}\right), \\
	L_i &= \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^\NN \Phi_+\left(2\ii;\NN,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{align}
for $i=0,\dots,\frac{\NN}{2}$, where
\begin{equation}
	\Phi_+(\mmf; \nnf, p) \coloneqq \sum\limits_{\dummy=0}^{\mmf/2} \binom{\nnf}{2\dummy} p^{2\dummy} (1-p)^{\nnf-2\dummy}.
\end{equation}
Conventionally, we set $(x_{-1}, L_{-1}) = (0,0)$.

The coordinates of any point on the Lorenz curve LHS can be expressed as convex sums of the neighbouring elbows,
\begin{align}
    x_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk}\right) x_{\ii-1} + p_{\ii\jj\kk} x_{\ii} \label{eq:lcsu_xcoord}\\
    L_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk} \right) L_{\ii-1} + p_{\ii\jj\kk} L_{\ii}, \label{eq:lcsu_lcoord}
\end{align}
where we use a three-index parametrisation for the coordinates, with $\ii$ determined by the count of the elbow that follows, $\jj$ by the number of Strange state copies and $\kk$ by the number of tensored maximally mixed state copies.
Therefore, for the initial state $\noisysn$ we have
\begin{equation}
	p_{\ii\jj\kk} = \frac{\jj}{8^{2\ii}\binom{\nn}{2\ii}},
\end{equation}
with $\ii=0,\dots,\nn,\ \jj(\ii) = 1,\dots,8^{2\ii}\binom{\nn}{2\ii}$ and $\kk = 1$.
For the target state $\noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{\mm}$,
\begin{equation}
	p_{\ii\jj\kk} = \frac{\kk + (\jj-1)9^{\mm}}{9^{\mm} 8^{2\ii}\binom{\tt}{2\ii}}
\end{equation}
with $\ii=0,\dots,\tt,\ \jj(\ii) = 1,\dots,8^{2\ii}\binom{\tt}{2\ii}$ and $\kk = 1,\dots,9^{\mm}$.

>>>>

\subsection{Strange state distillation in the unital fragment}\label{sec:lcsu}

We would now like to study the constraints coming solely from majorization in the unital fragment. 
We consider the problem of purifying $\NN$ copies of a noisy strange state $\noisysN$ into a smaller number of $\tt$ copies of a less noisy strange state $\rho(\epsilon')\too{\tt}$, with $\epsilon' \le \epsilon$ and $\tt \leq \nn$. 
Since the state $\frac{1}{3} \I$ is free in this fragment, the interconversion problem is unaffected if we tensor in copies of it. 
This turns out to simplify the Lorenz curve analysis, and so we focus on the process,
\begin{equation}\label{eq:sudist}
	\noisys\too{2\nn} \longrightarrow \noisys(\tt, \epsilon', \nn) \coloneqq  \noisys(\epsilon')\too{2\tt} \otimes \left( \frac{1}{3}\id \right)\too{2(n-t)},
\end{equation}
where all copies are even.
Since the initial state can be written in the form $\noisys(2\nn, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\rho}{\id/3}(x)$ for the generic noisy strange state $\noisys(\nn, \epsilon ,\tt)$.
The exact expressions for the coordinates of this Lorenz curve can take $2^3 = 8$ different forms, depending on whether the noise level $\epsilon$ is lower/higher than $3/7$, the parity of the number of copies $\NN$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
They are given in full detail in~\cref{app:lcsu_technical}.

Here we focus on the case of LHS points on Lorenz curves of even copies and noise levels $\epsilon \leq 3/7$.
The state $\rho(2\nn, \epsilon ,2\tt)$ has a Wigner distribution that takes on values $w_\ii$, with associated multiplicities $\mlt_\ii$, where
\begin{align}
w_\ii &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2\ii}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{2(\nn-\ii)} \\
m_\ii &= 8^{2\ii}\binom{2\nn}{2\ii}
\end{align}
From this, it is readily seen that $L_\sigma(x)$ has elbows on the LHS of its peak located at points $x=x_\ii$, where
\begin{equation}
x_\ii = \Phi_+ (\ii, \nn , \frac{8}{9}),
\end{equation}
with $\ii = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
\Phi_+(\ii, \nn, p) \coloneqq \sum_{\dummy=0}^{\ii} \binom{2\nn}{2\dummy} p ^{2\ii} (1-p)^{2(\nn-\ii)}.
\end{equation}
The value of the Lorenz curve at an elbow is given by
\begin{equation}
\lc{\rho}{\id/3}(x_{\ii}) \equiv L_{\ii} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{2\nn} \Phi_+\left(i;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

The coordinates of any point on the Lorenz curve LHS can be expressed as convex sums of the neighbouring elbows,
\begin{align}
    x_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk}\right) x_{\ii-1} + p_{\ii\jj\kk} x_{\ii} \label{eq:lcsu_xcoord}\\
    L_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk} \right) L_{\ii-1} + p_{\ii\jj\kk} L_{\ii}, \label{eq:lcsu_lcoord}
\end{align}
where we use a three-index parametrisation for the coordinates, with $\ii$ the index of the elbow that follows, $\jj = 1,\dots,\mlt_\ii$ and $\kk=1,\dots,9^{2(\nn-\tt)}$.
Therefore, for the initial state $\noisys\too{2\nn}$ we have
\begin{equation}
	p_{\ii\jj\kk} = \frac{\jj}{8^{2\ii}\binom{2\nn}{2\ii}},
\end{equation}
with $\ii=0,\dots,\nn,\ \jj(\ii) = 1,\dots,8^{2\ii}\binom{2\nn}{2\ii}$ and $\kk = 1$.
For the target state $\noisys(\epsilon')\too{2\tt} \otimes \left( \frac{1}{3}\id \right)\too{2(\nn-\tt)}$,
\begin{equation}
	p_{\ii'\jj'\kk'} = \frac{\kk' + (\jj'-1)9^{2(\nn-\tt)}}{9^{2(\nn-\tt)} 8^{2\ii'}\binom{2\tt}{2\ii'}}
\end{equation}
with $\ii'=0,\dots,\tt,\ \jj'(\ii') = 1,\dots,8^{2\ii'}\binom{\tt}{2\ii'}$ and $\kk' = 1,\dots,9^{2(\nn-\tt)}$.
